---
title: Ambulance Analysis
author: Rebecca Johnson
date: '`r format(Sys.Date(), "%B %d, %Y")`'
header-includes:
- \usepackage{float,booktabs,longtable,tabu,array}
- \usepackage[small]{caption}
- \captionsetup[table]{position=bottom}
- \floatplacement{figure}{H}  #make every figure with caption = h, this was the fix
- \floatplacement{table}{H}  #make every figure with caption = h, this was the fix
output:
pdf_document:
    fig_caption: yes
    fig_height: 9
    fig_width: 9
    latex_engine: xelatex
    keep_tex: true
    keep_md: true
    toc: true
geometry: "left=1in,right=1in,top=1in,bottom=1in"
graphics: yes
fontsize: 11pt
---

```{r, include=FALSE, echo=FALSE}
library(ggplot2)
library(dplyr)
library(stringr)
library(MASS)
library(lubridate)
library(stargazer)
library(reshape2)
library(xtable)
library(RColorBrewer)
library(scales)
library(here)

source(here("src", "R", "000_rmd_setup.R"))
source(here("src", "R", "000_constants.R"))
source(here("src", "R", "001_viz_utils.R"))
```

# 1. Data cleaning

```{r}
## load data
df_raw <- read.csv(file.path(INTERMEDIATE_DATA_DIR, "df_forambulanceuse.csv"))
```

```{r}
## different types of dispositions in safetypad
df_initial <- df_raw %>%
  mutate(
    femsresponse_char = gsub("\\s+", "", as.character(disposition)),
    FEMS_response = ifelse(femsresponse_char == "" &
      event_status %in% c(
        "NTL Handled - Canceled",
        "NTL Handled - Clinical Referral",
        "NTL Handled - RSC"
      ),
    "Not in SafetyPAD\n(canceled; clin. referral; or self care)",
    ifelse(femsresponse_char == "" &
      !event_status %in% c(
        "NTL Handled - Canceled",
        "NTL Handled - Clinical Referral",
        "NTL Handled - RSC"
      ),
    "Not in SafetyPAD\n(Other)",
    femsresponse_char
    )
    ),
    dispatched_ts = mdy_hm(dispatched)
  ) %>%
  dplyr::select(
    num_1, date, event_status, dispo_broad,
    FEMS_response, dispatched_ts
  ) %>%
  filter(dispo_broad %in% c("NTL control", "NTL treatment")) %>%
  rename(detailed_timestamp = dispatched_ts) %>%
  distinct()

# Output proportions of responses to file for supplemental appendix
tbl <- df_initial %>%
  filter(FEMS_response != "") %>%
  arrange(num_1, detailed_timestamp) %>%
  distinct(num_1, FEMS_response, .keep_all = TRUE) %>%
  group_by(num_1, dispo_broad) %>%
  summarize(
    all_responses = str_replace(
      paste(FEMS_response, collapse = "; "), "\n", " "
    ),
    .groups = "drop") %>%
  group_by(dispo_broad, all_responses) %>%
  mutate(numer = n()) %>%
  group_by(dispo_broad) %>%
  mutate(denom = n()) %>%
  group_by(dispo_broad, all_responses) %>%
  summarize(prop = numer / denom) %>%
  distinct() %>%
  filter(prop > 0.01) %>%
  arrange(dispo_broad, desc(prop)) %>%
  xtable() %>%
  print(file = file.path(TABLES_DIR, "recorded_responses.tex"))
```

```{r}
sprintf(
  "There are %s unique ids with %s appearances in the analytic sample",
  length(unique(df_initial$num_1)),
  nrow(df_initial)
)
```

```{r}
## see number of rows for each person
appearances_perid <- df_initial %>%
  group_by(num_1) %>%
  summarise(count_appearances = n()) %>%
  arrange(desc(count_appearances))

## status codes for appearances
type_appearances_perid <- df_initial %>%
  group_by(num_1) %>%
  arrange(detailed_timestamp) %>%
  summarise(all_appearances = paste(FEMS_response, collapse = "; "))


## look at high count id
# sprintf("Exploring why we have multiple appearances per id")
# df_initial %>% filter(num_1 == "F180183376")

## merge appearences per id onto main data
df_initial_wcountappear <- merge(df_initial,
  appearances_perid,
  on = "num_1",
  how = "left"
)

df_initial_wtypeappear <- merge(df_initial_wcountappear,
  type_appearances_perid,
  on = "num_1",
  how = "left"
)


## sort by time and get
## latset timestamp observation
df_analytic <- df_initial_wtypeappear %>%
  group_by(num_1) %>%
  arrange(desc(detailed_timestamp), .by_group = TRUE) %>%
  filter(row_number() == 1) %>%
  filter(!duplicated(num_1))


sprintf(
  "After taking the latest-time status for each id, there are %s unique ids with %s appearances",
  length(unique(df_analytic$num_1)),
  nrow(df_analytic)
)
```

```{r}
df_analytic %>%
  group_by(FEMS_response) %>%
  summarise(count = n()) %>%
  arrange(desc(count))

dispatch_codes <- setdiff(
  unique(df_analytic$FEMS_response),
  c(
    "Not in SafetyPAD\n(canceled; clin. referral; or self care)",
    "Not in SafetyPAD\n(Other)"
  )
)
```

```{r}
## two measures of ambulance use
df_analytic <- df_analytic %>%
  mutate(
    ambulanceuse_lastappearance = ifelse(FEMS_response == "Treated,TransportedbyEMS",
      1, 0
    ),
    ambulanceuse_anyappearance = ifelse(grepl("Treated,TransportedbyEMS", all_appearances),
      1, 0
    ),
    ambulancedispatch_anyappearance = ifelse(FEMS_response %in% dispatch_codes, 1, 0),
    ambulance_summary = case_when(
      ambulanceuse_lastappearance == 1 &
        ambulancedispatch_anyappearance == 1 ~
      "Amb. dispatched +\ntransported caller",
      ambulanceuse_lastappearance == 0 &
        ambulancedispatch_anyappearance == 1 ~ "Amb. dispatched but\nno transport",
      ambulanceuse_lastappearance == 0 &
        ambulancedispatch_anyappearance == 0 ~ "No amb. dispatch\nor transport",
      TRUE ~ "Amb. transport no dispatch"
    ),
    event_broad = case_when(
      grepl("Canceled", event_status) ~ "Cancelled",
      grepl("Referral", event_status) ~ "Clin. referral",
      grepl("RSC", event_status) ~ "Self-care",
      event_status == "Study reject" ~ "",
      TRUE ~ "Other"
    ),
    ambulance_wstatus = ifelse(ambulance_summary == "No dispatch or transport",
      sprintf("%s:\n%s", ambulance_summary, event_broad),
      ambulance_summary
    )
  )

prop.table(table(df_analytic$dispo_broad, df_analytic$ambulance_wstatus), margin = 1)
```

```{r}
## code dispatch no transport
prop_911response <- data.frame(prop.table(table(df_analytic$dispo_broad, df_analytic$ambulance_wstatus), margin = 1))
colnames(prop_911response) <- c("dispo_broad", "outcome_911", "proportion")
bottom_levels <- c(
  "Dispatch but no transport",
  "Dispatch and transport"
)
top <- c("No dispatch or transport:\nOther")
prop_911response <- prop_911response %>%
  filter(dispo_broad %in% c("NTL treatment", "NTL control")) %>%
  mutate(
    outcome_911_ordered = factor(outcome_911,
      levels = c(
        top, setdiff(
          unique(prop_911response$outcome_911),
          c(
            "Dispatch and transport",
            "Dispatch but no transport", top
          )
        ),
        bottom_levels
      ),
      ordered = TRUE
    ),
    percent = proportion * 100,
    percent_label = sprintf("%s%%", round(percent, 0))
  )


ggplot(prop_911response, aes(
  x = outcome_911_ordered, y = percent, fill = dispo_broad,
  group = dispo_broad
)) +
  geom_bar(stat = "identity", position = "dodge", color = "black") +
  theme_new(base_size = 18) +
  labs(fill = "") +
  xlab("Group") +
  ylab("Percent") +
  scale_fill_manual(values = c(
    "NTL treatment" = TREATMENT_COLOR,
    "NTL control" = CONTROL_COLOR
  )) +
  guides(fill = guide_legend(ncol = 2)) +
  theme(legend.position = "bottom") +
  ylab("Percent of group") +
  xlab("") +
  geom_label(aes(
    x = outcome_911_ordered, y = percent,
    label = round(percent, 2),
    group = dispo_broad
  ),
  fill = "white",
  position = position_dodge(width = 1)
  )

ggsave(file.path(FIGURE_DIR, "detailed_ambulance_perc.pdf"),
  plot = last_plot(),
  device = "pdf",
  width = 12,
  height = 8
)
```

```{r}
## among treatment callers, event statuses and correspondence
events_amb <- df_analytic %>%
  filter(dispo_broad == "NTL treatment") %>%
  mutate(event_medium = case_when(
    grepl("Canceled", event_status) ~ "Cancelled",
    grepl("Referral", event_status) ~ "Clin. referral",
    grepl("RSC", event_status) ~ "Self-care",
    grepl("ALS", event_status) ~ "ALS",
    grepl("BLS", event_status) ~ "BLS",
    TRUE ~ "Other"
  )) %>%
  group_by(event_medium, ambulance_summary) %>%
  summarise(count = n(), perc = (count / nrow(df_analytic %>% filter(dispo_broad == "NTL treatment"))) * 100) %>%
  ungroup() %>%
  mutate(event_ordered = factor(event_medium,
    levels = c(
      "ALS",
      "BLS",
      "Cancelled",
      "Clin. referral",
      "Self-care",
      "Other"
    ),
    ordered = TRUE
  ))
```

```{r}
ggplot(events_amb, aes(x = event_ordered, y = perc, fill = ambulance_summary)) +
  geom_bar(stat = "identity", position = "dodge", color = "black") +
  theme_new(base_size = 16) +
  theme(legend.position = c(0.7, 0.7)) +
  xlab("Call response code") +
  ylab("Percent of treatment group\n(sums to 100 across all bars)") +
  scale_fill_manual(values = c(
    "#272932", "#427EA8",
    "#B6C2D9"
  )) +
  labs(fill = "")

ggsave(file.path(FIGURE_DIR, "txgroup_callresponse_vamb.pdf"),
  plot = last_plot(),
  device = "pdf",
  width = 12,
  height = 8
)
```

```{r}
## write file for use in updated Medicaid dx analysis
write.csv(df_analytic, file.path(INTERMEDIATE_DATA_DIR, "callresponse_forposttx.csv"),
  row.names = FALSE
)
```

# 2. Ambulance use

## 2.1 Descriptive across entire study period

```{r}
prop_ambulance_last <- df_analytic %>%
  group_by(dispo_broad, ambulanceuse_lastappearance) %>%
  summarise(count_used = n()) %>%
  left_join(df_analytic %>% group_by(dispo_broad) %>% summarise(total_count = n())) %>%
  mutate(prop_use_last = count_used / total_count) %>%
  dplyr::select(
    dispo_broad, ambulanceuse_lastappearance, prop_use_last, count_used,
    total_count
  ) %>%
  rename(ambulanceuse = ambulanceuse_lastappearance)
prop_ambulance_any <- df_analytic %>%
  group_by(dispo_broad, ambulanceuse_anyappearance) %>%
  summarise(count_used = n()) %>%
  left_join(df_analytic %>% group_by(dispo_broad) %>% summarise(total_count = n())) %>%
  mutate(prop_use_any = count_used / total_count) %>%
  dplyr::select(dispo_broad, ambulanceuse_anyappearance, prop_use_any) %>%
  rename(ambulanceuse = ambulanceuse_anyappearance)


prop_ambulance_either <- merge(prop_ambulance_last, prop_ambulance_any,
  by = c("dispo_broad", "ambulanceuse")
) %>%
  mutate(ambulanceuse_descriptive = factor(ambulanceuse,
    levels = c(0, 1),
    labels = c(
      "No ambulance\ntransport for call",
      "Ambulance\ntransport for call"
    )
  ))


prop_ambulance_last %>%
  mutate(ambulance_transport = ifelse(ambulanceuse == 0,
    "No transport",
    "Transport"
  )) %>%
  dplyr::select(dispo_broad, ambulance_transport, count_used, total_count)
```

```{r}
## similar for dispatch
prop_dispatch_any <- df_analytic %>%
  group_by(dispo_broad, ambulancedispatch_anyappearance) %>%
  summarise(count_used = n()) %>%
  left_join(df_analytic %>% group_by(dispo_broad) %>% summarise(total_count = n())) %>%
  mutate(prop_use_any = count_used / total_count) %>%
  dplyr::select(dispo_broad, ambulancedispatch_anyappearance, prop_use_any) %>%
  rename(ambulancedispatch = ambulancedispatch_anyappearance) %>%
  mutate(ambulancedispatch_descriptive = factor(ambulancedispatch,
    levels = c(0, 1),
    labels = c(
      "No ambulance\ndispatch for call",
      "Ambulance\ndispatch for call"
    )
  ))
```

```{r}

## Main text figure on prop by group
ggplot(prop_ambulance_either, aes(
  x = ambulanceuse_descriptive, y = prop_use_last,
  fill = dispo_broad, group = dispo_broad
)) +
  geom_bar(stat = "identity", position = "dodge") +
  xlab("") +
  ylab("Proportion of each group") +
  labs(fill = "") +
  scale_fill_manual(values = c(
    "NTL treatment" = TREATMENT_COLOR,
    "NTL control" = CONTROL_COLOR
  )) +
  geom_label(aes(x = ambulanceuse_descriptive, y = prop_use_last, label = round(prop_use_last, 2)),
    position = position_dodge(width = 1), size = 6, fill = "white"
  ) +
  theme_new(base_size = 24) +
  theme(
    legend.position = c(0.3, 0.9),
    legend.background = element_blank()
  )

## save results
ggsave(file.path(FIGURE_DIR, "propambulance_laststatus.pdf"),
  plot = last_plot(),
  device = "pdf",
  width = 12,
  height = 8
)
```

```{r}
## Appendix figure on prop by group
ggplot(prop_ambulance_either, aes(
  x = ambulanceuse_descriptive, y = prop_use_any,
  fill = dispo_broad, group = dispo_broad
)) +
  geom_bar(stat = "identity", position = "dodge", alpha = 0.8) +
  xlab("") +
  ylab("Proportion of each group") +
  labs(fill = "") +
  scale_fill_manual(values = c(
    "NTL treatment" = TREATMENT_COLOR,
    "NTL control" = CONTROL_COLOR
  )) +
  geom_label(aes(x = ambulanceuse_descriptive, y = prop_use_any, label = round(prop_use_any, 2)),
    position = position_dodge(width = 1), size = 6, fill = "white"
  ) +
  theme_new(base_size = 24) +
  theme(
    legend.position = c(0.3, 0.9),
    legend.background = element_blank()
  )

## save results
ggsave(file.path(FIGURE_DIR, "propambulance_any.pdf"),
  plot = last_plot(),
  device = "pdf",
  width = 12,
  height = 8
)
```

```{r}
## Main text figure on prop by group
ggplot(prop_dispatch_any, aes(
  x = ambulancedispatch_descriptive, y = prop_use_any,
  fill = dispo_broad, group = dispo_broad
)) +
  geom_bar(stat = "identity", position = "dodge", alpha = 0.8) +
  xlab("") +
  ylab("Proportion of each group") +
  labs(fill = "") +
  scale_fill_manual(values = c(
    "NTL treatment" = TREATMENT_COLOR,
    "NTL control" = CONTROL_COLOR
  )) +
  geom_label(aes(x = ambulancedispatch_descriptive, y = prop_use_any, label = round(prop_use_any, 2)),
    position = position_dodge(width = 1), size = 6, fill = "white"
  ) +
  theme_new(base_size = 24) +
  theme(
    legend.position = c(0.3, 0.9),
    legend.background = element_blank()
  )

## save results
ggsave(file.path(FIGURE_DIR, "propamb_dispatch.pdf"),
  plot = last_plot(),
  device = "pdf",
  width = 12,
  height = 8
)
```

## 2.2 regressions across entire study period

Language on controls from PAP:

If baseline balance analysis justifies their inclusion we will include pre-treatment covariates for NTL-eligible callers who have a 911 calling history, and indicators for the day of the pilot period.

```{r}
run_regs <- function(data, outcome_varname, treatment_varname = "treatment",
                     include_dayFE = FALSE, dayFE_name) {

  ##
  if (include_dayFE == FALSE) {
    reg_formula <- formula(sprintf("%s ~ %s", outcome_varname, treatment_varname))
    reg_results <- glm(reg_formula, data = data, family = "binomial")
    return(reg_results)
  } else {
    reg_formula <- formula(sprintf(
      "%s ~ %s + factor(%s)", outcome_varname, treatment_varname,
      dayFE_name
    ))
    reg_results <- glm(reg_formula, data = data, family = "binomial")
    return(reg_results)
  }
}
```

```{r}
df_analytic <- df_analytic %>%
  mutate(
    treatment = ifelse(dispo_broad == "NTL treatment", 1, 0),
    day_rounded = as.Date(date, format = "%Y-%m-%d")
  )

ambulancemain_nocontrols <- run_regs(data = df_analytic, outcome_varname = "ambulanceuse_lastappearance")
ambulancerobust_nocontrols <- run_regs(data = df_analytic, outcome_varname = "ambulanceuse_anyappearance")
ambulancedispatch_nocontrols <- run_regs(data = df_analytic, outcome_varname = "ambulancedispatch_anyappearance")
```

```{r}
stargazer(
  ambulancemain_nocontrols,
  ambulancerobust_nocontrols,
  ambulancedispatch_nocontrols,
  title = "Effect of Nurse Triage Line on Ambulance Dispatches and Transport (logistic regression)",
  label = "tab:amb_full_results",
  dep.var.labels = c("Amb. transport (last status)", "Amb. transport (any status)", "Amb. dispatch (any status)"),
  covariate.labels = c("Treatment", "Constant"),
  omit = "day_rounded",
  font.size = "scriptsize",
  digits = 3,
  out = file.path(TABLES_DIR, "regression_ambulance_main.tex")
)
```

## 2.3 Visualizing

Not using predicted probabilities code for now

For now, showing raw means and p-value

```{r}
prop_forgraph <- df_analytic %>%
  group_by(treatment) %>%
  summarise(
    transport = mean(ambulanceuse_lastappearance),
    dispatch = mean(ambulancedispatch_anyappearance)
  ) %>%
  ungroup() %>%
  mutate(treatment_descriptive = ifelse(treatment == 1,
    "NTL treatment",
    "NTL control"
  )) %>%
  dplyr::select(-treatment) %>%
  melt(, id.vars = "treatment_descriptive") %>%
  mutate(
    variable_descriptive = ifelse(variable == "transport",
      "Amb. transport\n",
      "Amb. dispatch\n(includes transport)"
    ),
    perc = value * 100
  )

ggplot(prop_forgraph, aes(
  x = variable_descriptive, y = perc,
  group = treatment_descriptive, fill = treatment_descriptive
)) +
  geom_bar(stat = "identity", position = "dodge") +
  theme_new(base_size = 24) +
  scale_fill_manual(values = c(
    "NTL treatment" = TREATMENT_COLOR,
    "NTL control" = CONTROL_COLOR
  )) +
  labs(fill = "") +
  theme(
    legend.position = c(0.75, 0.9),
    legend.background = element_blank()
  ) +
  xlab("") +
  ylab("Percentage of group") +
  scale_y_continuous(breaks = pretty_breaks(n = 10)) +
  annotate("text", x = c(1.2, 2.2), y = c(60, 48), label = "p < 0.001", size = 6)

ggsave(file.path(FIGURE_DIR, "ambuse_maintext.pdf"),
  plot = last_plot(),
  device = "pdf",
  width = 12,
  height = 8
)
```

## 2.4. Regressions separately for each month

```{r}
df_analytic <- df_analytic %>%
  mutate(month = format(as.Date(date), "%Y-%m"))
```

```{r}
reg_bymonth <- function(data, month_name) {
  filtered_data <- data %>% filter(month == month_name)
  reg_results <- run_regs(filtered_data, outcome_varname = "ambulanceuse_lastappearance")
  return(reg_results)
}
```

```{r}
all_months <- unique(df_analytic$month)


regs_allmonths <- lapply(all_months, reg_bymonth, data = df_analytic)
```

```{r}
summary(regs_allmonths[[1]])$coefficients
```

```{r}
treatment_effect <- lapply(regs_allmonths, function(x) coefficients(x)["treatment"])
se <- lapply(regs_allmonths, function(x) summary(x)$coefficients["treatment", 2])
pval <- lapply(regs_allmonths, function(x) summary(x)$coefficients["treatment", 4])
summarize_estimates_bymonth <- data.frame(
  month = all_months, treatment_effect = unlist(treatment_effect),
  se = unlist(se), pval_uncorrected = unlist(pval)
) %>%
  arrange(month) %>%
  mutate(
    lower = treatment_effect - 1.96 * se,
    upper = treatment_effect + 1.96 * se
  )
summarize_estimates_bymonth

ggplot(summarize_estimates_bymonth, aes(x = factor(month), y = treatment_effect)) +
  geom_point() +
  geom_errorbar(aes(x = factor(month), ymin = lower, ymax = upper),
    width = 0.5
  ) +
  theme_new(base_size = 24) +
  ylim(-3, 0.1) +
  coord_flip() +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  xlab("Month") +
  ylab("Treatment effect on call likelihood\nof ambulance use\n(negative: treatment group\nhad lower use)")

ggsave(file.path(FIGURE_DIR, "coef_amb_bymonth.pdf"),
  plot = last_plot(),
  device = "pdf",
  width = 12,
  height = 8
)
```

```{r}
pvals_corrected <- p.adjust(summarize_estimates_bymonth$pval_uncorrected,
  method = "bonferroni",
  n = length(summarize_estimates_bymonth$pval_uncorrected)
)

summarize_estimates_bymonth$pvals_corrected <- pvals_corrected

summarize_estimates_bymonth %>%
  dplyr::select(month, treatment_effect, pval_uncorrected, pvals_corrected) %>%
  rename(
    Month = month,
    `Treatment effect` = treatment_effect,
    `p-value (uncorrected)` = pval_uncorrected,
    `p-value (corrected)` = pvals_corrected
  ) %>%
  xtable(
    caption = "Treatment effects on ambulance use by month: coefficients, p values, and Bonferonni-corrected p values",
    label = "tab:logit_bymonth",
    # align = c("l", "r", "r", "r"),
    digits = c(3, 4, 4, 10, 10)
  ) %>%
  print(
    include.rownames = FALSE,
    caption.placement = "top",
    file = file.path(TABLES_DIR, "bonferonni_pvalues.tex")
  )

# print(xtable(summarize_estimates_bymonth %>% dplyr::select(
#   month, treatment_effect, pval_uncorrected,
#   pvals_corrected
# ),
# digits = c(3, 4, 4, 10, 10)
# ),
# include.rownames = FALSE
# )
```

